---
title: "TAC linearity summary"
author: "Kelsey Jesser"
date: "11/12/22"
output: html_document
---
#clear R env
```{r message=FALSE, warning=FALSE}
rm(list = ls())
```

Load libraries
```{r message=FALSE, warning=FALSE}
library(ggplot2)
library(ggpmisc)
library(plyr)
library(dplyr)
library(tidyr)
```

Import data file
```{r message=FALSE, warning=FALSE}
TAC<-read.csv("Sample_data.csv", na.strings=c("","NA"))
```

Plot a standard curve and summarize efficiency for each target
```{r message=FALSE, warning=FALSE}
#Format data and add column with logs of quantities
TAC$CT<-as.numeric(gsub(",","",TAC$CT))

TAC$Quantity<-as.numeric(gsub(",","",TAC$Quantity))
TAC$Quantity_log<-log10(TAC$Quantity)

#summarize efficiencies
TAC<-TAC[!is.na(TAC$CT),]
TAC<-TAC[!is.na(TAC$Quantity_log),]
TAC<-TAC[rowSums(is.na(TAC)) != ncol(TAC), ]
efficiency_summ<-ddply(TAC, "Target.Name", function(x) {
  model <- lm(CT ~ Quantity_log, data = x)
  coef(model)})

colnames(efficiency_summ) <- c("Target.Name", "Y.intercept","Slope")

efficiency_summ$Efficiency<-10^(-1/efficiency_summ$Slope)-1

#create a list of dataframes by target
targets=by(TAC, TAC[,"Target.Name"], function(x) x)

#plot of standard curve points across all targets
p<-ggplot(TAC, aes(x = Quantity_log, y = CT)) +
  geom_point()
p

#plot standard curves for all targets
target_plots <- function(targets){
  ggplot(targets, aes(x = Quantity_log, y = CT)) +
    geom_point()+
    stat_poly_line(formula = y~x, se=FALSE) +
    stat_poly_eq(aes(label = paste(after_stat(eq.label), after_stat(rr.label), sep = "*\", \"*")), formula= y~x, label.y=0.25)
  }

q1 <- lapply(targets, target_plots) 

q2 <- lapply(seq_along(q1), function(i) { 
  q1[[i]] + ggtitle(names(targets)[i])
})

q2 #careful--this command makes a lot of plots!

#faceted plot with all adjusted standard curves for all targets
r<-ggplot(TAC, aes(x = Quantity_log, y = CT)) +
  geom_point() +
  facet_wrap(~ Target.Name)+
  stat_poly_line(formula = y~x, se=FALSE) +
  stat_poly_eq(aes(label = paste(after_stat(eq.label), after_stat(rr.label), sep = "*\", \"*")), formula= y~x, label.y=20)
r
```

Plot standard curves and summarize efficiencies without last point in the curve (where it flattens out as it approaches the LOD)
```{r message=FALSE, warning=FALSE}
#remove last point on standard curve
TAC_adj <- TAC[TAC$Quantity_log != 0, ]
#TAC_adj <- TAC[TAC$Quantity_log != 0, ] #run this to look at standard curves without last two dilutions to improve efficencies of AMR gene assays and other targets with lower LODs

#create a list of dataframes by target
targets_adj=by(TAC_adj, TAC_adj[,"Target.Name"], function(x) x)

#plot adjusted standard curves for all targets
target_plots_adj <- function(targets_adj){
  ggplot(targets_adj, aes(x = Quantity_log, y = CT)) +
    geom_point()+
    stat_poly_line(formula = y~x, se=FALSE) +
    stat_poly_eq(aes(label = paste(after_stat(eq.label), after_stat(rr.label), sep = "*\", \"*")), formula= y~x, label.y=0.25)
  }

s1 <- lapply(targets_adj, target_plots_adj) 

s2 <- lapply(seq_along(s1), function(i) { 
  s1[[i]] + ggtitle(names(targets_adj)[i])
})

s2

#faceted plot with all adjusted standard curves for all targets
t<-ggplot(TAC_adj, aes(x = Quantity_log, y = CT)) +
  geom_point() +
  facet_wrap(~ Target.Name)+
  stat_poly_line(formula = y~x, se=FALSE) +
  stat_poly_eq(aes(label = paste(after_stat(eq.label), after_stat(rr.label), sep = "*\", \"*")), formula= y~x, label.y=20)
t

```

Summarize dilutions with detection and high and low CT values for each target FIX THIS
```{r message=FALSE, warning=FALSE}
TAC_CT<-TAC %>% select(Target.Name, Sample.Name, CT, R.superscript.2.)

high_low<-TAC_CT %>% 
  group_by(Target.Name) %>%
  mutate(
    CT_min=min(CT, na.rm=T),
    CT_max=max(CT, na.rm=T)) %>%
  mutate(
    Dilution_min=min(Sample.Name, na.rm=T), 
    Dilution_max=max(Sample.Name, na.rm=T)) %>%
  select(-Sample.Name) %>%
  select(-CT)
```

Summarize adjusted (without lowest dilution) slope, efficiency, and R2 values for each target 
```{r message=FALSE, warning=FALSE}
R2<- TAC_adj %>%
   group_by(Target.Name) %>%
   dplyr::summarise(model = list(lm(CT ~ Quantity_log, data = cur_data())),
             coef = list(coef(model[[1]])),
             R2_adj = summary(model[[1]])$r.sq)%>%
   unnest_wider(coef, names_repair = 'unique')

R2$Efficiency_adj<- 10^(-1/R2$Quantity_log)-1
```

Summarize NTC results
```{r message=FALSE, warning=FALSE}
TAC_NTC<-read.csv("sample_data.csv", na.strings=c("","NA"))

TAC_NTC <- TAC_NTC %>%
  subset(Sample.Name=="NTC")

TAC_NTC$CT<-as.numeric(gsub(",","",TAC_NTC$CT))
TAC_NTC<- TAC_NTC%>% 
  select(Target.Name, CT)%>%
  dplyr::rename(NTC_CT=CT)

```

Summary table
```{r message=FALSE, warning=FALSE}
summary<-merge(efficiency_summ, R2)
summary<-merge(high_low, summary)
summary<-merge(TAC_NTC, summary)
summary <- summary %>% 
  distinct(.keep_all = TRUE) %>%
  select(-model) %>%
  relocate(R.superscript.2., .before=Efficiency) %>%
  relocate(NTC_CT, .before=Y.intercept)%>%
  dplyr::rename(R2=R.superscript.2.) %>%
  dplyr::rename(Y.intercept_adj='(Intercept)') %>%
  dplyr::rename(Slope_adj=Quantity_log)%>%
  mutate_if(is.numeric, round, digits=2)

write.csv(summary,file="TAC_sample_linearity_summary.csv")
```
